---
title: "A Report On Lake Ontario's Microbes"
author: "Jordan"
date: "2025-02-19"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#install.packages("tinytex")
#tinytex::install_tinytex()
knitr::opts_chunk$set(echo = TRUE)
```

#Prepre the environment

```{r packages}
#Load Libraries 
library(tidyverse)
```

# Load in the Lake Ontario data

```{r load-data}
#Load in Lake Ontario microbial community data
sample_and_taxon <-
 read_csv("data/sample_and_taxon.csv")

#Inspct data
glimpse(sample_and_taxon)
```

# Lake Ontario microbial abundance vs temperature

```{r cells_vs_temp}
ggplot(data = sample_and_taxon, aes(x = temperature, y = cells_per_ml))+
  labs(x = "Temperature (C)", y = "Cell Abundance (cells/mL)") +
  geom_point(aes(color = env_group)) +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() +
  scale_color_manual(values = c("cornflowerblue", "goldenrod", "firebrick")) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

The above plot shows that:

- There's a positive relationship between temperature and cell abundances.
- For example, deep samples are the coldest and have the fewest cells.

```{r data-summary, include=FALSE}
#How many samples do I have?
n_samples <-
  sample_and_taxon %>%
  #Only pick out sample id column
  select(sample_id) %>%
  #Number of distinct names in the column
  n_distinct()

#Calculate min and max temps
#min
min_temp <-
  sample_and_taxon %>%
  summarize(round(min(temperature),digits = 1)) %>%
  pull()
#max
max_temp <-
  sample_and_taxon %>%
  summarize(round(max(temperature),digits = 1)) %>%
  pull()
```

The total number of samples is `r n_samples`. For this set of samples, temperature ranged from a minimum of `r min_temp` Celsius to a maximum of `r max_temp` Celsius.
